---
title: "Team file"
author: "Sonia Thomas and John Nownes"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(readxl)
library(reshape2)
library(dplyr)
library(tidyr)
library(ggplot2)
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
bio<- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
```

## Lab 4

### Part one: Cleaning the Data

#### Questions 1-3

```{r }
defClean<- defense %>% 
  mutate(Name =as.factor(Name)) %>%
  mutate(Opponent_Opponent = as.factor(Opponent_Opponent)) %>%
  mutate_at(vars(contains("Tackles")), funs(as.numeric)) %>%
  mutate_at(vars(contains("Turnover")), funs(as.numeric)) %>%
  mutate_at(vars(contains("Pass")), funs(as.numeric)) 
  
str(defClean)

offClean<- offense %>% 
  mutate(Name =as.factor(Name)) %>%
  mutate(Opponent_Opponent = as.factor(Opponent_Opponent)) %>%
  mutate_at(vars(contains("Rushing")), funs(as.numeric)) %>%
  mutate_at(vars(contains("Receiving")), funs(as.numeric)) %>%
  mutate_at(vars(contains("Passing")), funs(as.numeric)) 
 
str(offClean)

bioClean<- bio %>% 
  mutate(Name =as.factor(Name)) %>%
  mutate_at(vars(contains("Weight")), funs(as.numeric)) %>%
  separate(Height, into = c("ft", "inches")) %>%
  mutate(Height = as.numeric(ft)*12 + as.numeric(inches)) %>%
  select(-c(ft, inches)) %>%
  select(1:2, Height, everything())
  
str(bioClean)
```

### Part two: Tidying (1)

1.
    ```{r}
    defClean2 = defClean %>%
      pivot_longer(Tackles_Solo:Pass_PB, names_to = 'Statistic')
    ```
    
2. 
    ```{r}
    defClean2 %>%
      ggplot(aes(value)) + geom_histogram(binwidth = .5) + facet_wrap(~Statistic) + ggtitle("Defensive Statistic Histograms")
    ```
    * Looking at the above plot of all the defensive histograms, it is clear that the three turnover categories (forced fumble, fumble recovery, and interceptions) are the most rare defensive statistics. Sacks are also quite rare as well.
    
3.
    ```{r}
    dat2_3 = defClean2 %>%
  filter(Opponent_Opponent == "West Virginia" | Opponent_Opponent == "Kansas State", Statistic == "Tackles_Solo") %>%
  pivot_wider(names_from = Opponent_Opponent, values_from =  value)

dat2_3$`West Virginia`[is.na(dat2_3$`West Virginia`)] = 0
dat2_3$`Kansas State`[is.na(dat2_3$`Kansas State`)] = 0
dat2_3 = dat2_3 %>%
  filter(`West Virginia` != 0 | `Kansas State` != 0)

p = dat2_3 %>%
  ggplot(aes(x = `West Virginia`, y = `Kansas State`)) + geom_count() + coord_cartesian(xlim = c(0,6))
p = p + ggtitle("Scatter Plot of Solo Tackles For Each Player\n Against West Virginia and Kansas State")
p
    ```
    * Iowa State had a better defense against Kansas State, based on the comparison of the total solo tackles against each team. From the plot above, you can tell that most individual players recorded a higher number of solo tackles against Kansas State than they did against West Virginia. In addition, by summing the total number of solo tackles against each team, we see that ISU tallied 45 total solo tackles against Kansas State and only 34 total solo tackles against West Virginia.
    
### Part two: Tidying (2)

1.
    ```{r}
    bioClean <- bioClean %>%
      separate(col=Hometown, into=c('City', 'State'), sep=',') 

    head(bioClean, n=3)
    ```
    
2.
    ```{r}
    table(bioClean$State)
    ```
    * The output above you can see how many players are from each state. You can see that most players are from Iowa. 
    
### Part three: Joining data frames

1.
    ```{r}
    
    ```
    
2.
    ```{r}
    
    ```
    
3.
    ```{r}
defOff<- full_join(offClean, defClean, by= c('Opponent_Opponent', 'Name'))
defOff2<- select(defOff, -c('Passing_CMP-ATT')) %>% 
  distinct(defOff, Name, Opponent_Opponent, .keep_all = TRUE)

avgISU<- defOff2 %>%
  mutate(Rushing_ATT = sum(Rushing_ATT, na.rm = TRUE)) %>%
  mutate(Rushing_YDS = sum(Rushing_YDS, na.rm = TRUE)) %>%
  mutate(Rushing_TD = sum(Rushing_TD, na.rm = TRUE)) %>%
  mutate(Receiving_REC = sum(Receiving_REC, na.rm = TRUE)) %>%
  mutate(Receiving_TD = sum(Receiving_TD, na.rm = TRUE)) %>%
  mutate(Passing_YDS = sum(Passing_YDS, na.rm = TRUE)) %>%
  mutate(Passing_TD = sum(Passing_TD, na.rm = TRUE)) %>%
  mutate(Passing_INT = sum(Passing_INT, na.rm = TRUE)) %>%
  mutate(Tackles_Solo = sum(Tackles_Solo, na.rm = TRUE)) %>%
  mutate(Tackles_ASST = sum(Tackles_ASST, na.rm = TRUE)) %>%
  mutate(Tackles_TFL = sum(Tackles_TFL, na.rm = TRUE)) %>%
  mutate(Tackles_Sack = sum(Tackles_Sack, na.rm = TRUE)) %>%
  mutate(Turnover_FF = sum(Turnover_FF, na.rm = TRUE)) %>%
  mutate(Turnover_FR = sum(Turnover_FR, na.rm = TRUE)) %>%
  mutate(Turnover_INT = sum(Turnover_INT, na.rm = TRUE)) %>%
  mutate(Pass_QBH = sum(Pass_QBH, na.rm = TRUE)) %>%
  mutate(Pass_PB = sum(Pass_PB, na.rm = TRUE))
  
new<- head(avgISU, 1)
new1<- select(new, -c(Name, Opponent_Opponent))
new2<- new1 %>%
  pivot_longer(1:18, names_to = 'Type', values_to = 'Statistics')
new3<- new2 %>%
  mutate(Name = "ISU TEAM")

purdyB<- select(defOff, -c('Passing_CMP-ATT')) %>%
  filter(Name %in% c('Purdy, Brock')) %>%
  mutate(Rushing_ATT = sum(Rushing_ATT, na.rm = TRUE)) %>%
  mutate(Rushing_YDS = sum(Rushing_YDS, na.rm = TRUE)) %>%
  mutate(Rushing_TD = sum(Rushing_TD, na.rm = TRUE)) %>%
  mutate(Receiving_REC = sum(Receiving_REC, na.rm = TRUE)) %>%
  mutate(Receiving_TD = sum(Receiving_TD, na.rm = TRUE)) %>%
  mutate(Passing_YDS = sum(Passing_YDS, na.rm = TRUE)) %>%
  mutate(Passing_TD = sum(Passing_TD, na.rm = TRUE)) %>%
  mutate(Passing_INT = sum(Passing_INT, na.rm = TRUE)) %>%
  mutate(Tackles_Solo = sum(Tackles_Solo, na.rm = TRUE)) %>%
  mutate(Tackles_ASST = sum(Tackles_ASST, na.rm = TRUE)) %>%
  mutate(Tackles_TFL = sum(Tackles_TFL, na.rm = TRUE)) %>%
  mutate(Tackles_Sack = sum(Tackles_Sack, na.rm = TRUE)) %>%
  mutate(Turnover_FF = sum(Turnover_FF, na.rm = TRUE)) %>%
  mutate(Turnover_FR = sum(Turnover_FR, na.rm = TRUE)) %>%
  mutate(Turnover_INT = sum(Turnover_INT, na.rm = TRUE)) %>%
  mutate(Pass_QBH = sum(Pass_QBH, na.rm = TRUE)) %>%
  mutate(Pass_PB = sum(Pass_PB, na.rm = TRUE))

purdyB1<- head(purdyB, 1)
purdyB2<- select(purdyB1, -c(Opponent_Opponent))
purdyB3<- purdyB2 %>%
  pivot_longer(2:19, names_to = 'Type', values_to = 'Statistics')



completeISU<- full_join(new3, purdyB3, by=c('Type', 'Name', 'Statistics'))

completeISU

ggplot(data=completeISU, aes(x=Type, y = Statistics, fill = Name)) + geom_bar(stat='identity', position = position_dodge()) + theme(axis.text.x = element_text(angle =45, vjust = 0.5)) + xlab("Statistic") + ylab('Count') + ggtitle("Purdy's vs Team Performance ")
    ```
    * Based on output above you can see how the teams performance compares to Purdys performance. I would say that in most areas that when Purdy performs well in his respective statistic (all passing stats) the team peforms well too.
4.
    ```{r}
#Question: What statistic do redshirt freshman excel in? 
age<- full_join(avgISU, bioClean, by= c('Name'))

redfreshman<- age %>%
  filter(Class %in% c('Redshirt Freshman')) 
    
redfreshman1<- select(redfreshman, -c(Name, Opponent_Opponent))
redfreshman2<- redfreshman1 %>%
  pivot_longer(1:18, names_to = 'Type', values_to = 'Statistics')
redfreshman3<- redfreshman2 %>%
  mutate(Name = "REDSHIRT FRESHMAN")

ggplot(data=redfreshman3, aes(x=Type, y = Statistics, fill = Name)) + geom_bar(stat='identity', position = position_dodge()) + theme(axis.text.x = element_text(angle =45, vjust = 0.5)) + xlab("Statistic") + ylab('Count') + ggtitle("Redshirt freshman performance")
    ```
    * Redshirt freshman based on the output above seem to perform the best in the amount of passing touchdowns received. 